\chapter{Project Management}

Project management for Jan-Sunwai AI followed an iterative delivery approach with explicit weekly milestones, acceptance checkpoints, and operational hardening gates. The planning strategy balanced feature development with risk reduction, especially in AI reliability, security posture, and deployment reproducibility.

\section{Feasibility Study}

\subsection{Technical Feasibility}

The selected stack (React, FastAPI, MongoDB, Ollama) was technically feasible for local-first civic deployment due to mature ecosystem support, strong documentation, and compatibility with available development infrastructure. Key feasibility observations were:

\begin{itemize}
\item FastAPI provided rapid API development, typed schemas, and OpenAPI generation.
\item MongoDB supported flexible document structures suited to complaint histories, comments, and embedded metadata.
\item React enabled role-specific UI composition with route guards and dashboard specialization.
\item Ollama provided local model execution with configurable model tiers.
\item Docker Compose enabled reproducible environment orchestration across development and production-like setups.
\end{itemize}

A major early technical risk was model reliability under constrained VRAM. This was mitigated by introducing a staged inference pipeline and avoiding simultaneous heavy model loading.

\subsection{Time Schedule Feasibility}

The schedule was structured over 18 weeks, progressing from foundation to hardening and final submission. Work cadence used Monday--Saturday windows with planned Sunday off-days. Alternate Saturday breaks were also reserved. This realistic scheduling improved sustainability while preserving delivery continuity.

\subsection{Operational Feasibility}

Operational feasibility was validated by mapping municipal personas into role-based workflows:

\begin{itemize}
\item Citizen: complaint intake and tracking.
\item Worker: assignment execution and completion updates.
\item Department Head: queue supervision and status transitions.
\item Admin: approvals, triage governance, analytics, and exports.
\end{itemize}

The resulting workflow matched expected governance practices where supervisory roles require both visibility and controlled override capabilities.

\subsection{Implementation Feasibility}

Implementation was feasible under both manual and scripted setup modes. Automation scripts were created for Windows and Linux, and production deployment was validated through containerized orchestration. Environment configuration was standardized via a single backend environment file to reduce drift and support repeatable handover.

\section{Project Planning}

\subsection{Development Approach and Justification}

The project used an incremental and verification-driven approach with the following cycle:

\begin{enumerate}
\item Build baseline functionality for a module.
\item Validate through unit and integration checks.
\item Expose role workflow in UI and test end-to-end behavior.
\item Add security/performance hardening after functional stabilization.
\item Document operational procedures and release notes.
\end{enumerate}

This approach was chosen over a strict waterfall model because AI behavior and UX patterns required iterative calibration.

\subsection{Milestones and Deliverables}

The following longtable consolidates the key details for this subsection in a structured and comparable format.


\begin{longtable}{|p{0.12\textwidth}|p{0.24\textwidth}|p{0.43\textwidth}|p{0.13\textwidth}|}
\caption{Milestone Summary and Deliverables}\\
\toprule
\textbf{Window} & \textbf{Theme} & \textbf{Key Deliverables} & \textbf{Status} \\ \hline
\midrule
\endfirsthead
\toprule
\textbf{Window} & \textbf{Theme} & \textbf{Key Deliverables} & \textbf{Status} \\ \hline
\midrule
\endhead
Weeks 1--2 & Foundation & Repo setup, environment bootstrap, backend skeleton, DB connectivity, auth schema, complaint schema, initial CRUD and upload service & Completed \\ \hline
Weeks 3--5 & AI Pipeline & Model research, classifier iterations, geotagging, reverse geocoding, generation service, analyze endpoint integration & Completed \\ \hline
Week 6 & Frontend Base & Routing, upload flow, analyze hook, result view, integration with backend auth and analyze pipeline & Completed \\ \hline
Week 7 & Flow Closure & Map marker improvements, local draft persistence, error boundary, 404 route, full persona walkthrough & Completed \\ \hline
Week 8 & Security Layer & Upload hardening, rate limiting, sanitization, CORS lockdown, security headers, index hardening & Completed \\ \hline
Week 9 & NDMC Feature Pack & Worker panel, assignment engine, audit trail, analytics/heatmap, SLA badges, bulk actions/exports & Completed \\ \hline
Week 10 & Notification and Communication & Navbar badge, status-triggered notifications, email stub, forgot/reset password, profile editing, E2E notification chain test & Completed \\ \hline
Week 11 & Performance and Reliability & Image compression, lazy loading, graceful 503 degrade, profiling, resilience tests & Completed \\ \hline
Week 12 & Production Hardening & Docker production files, compose profile, env plan, API versioning, backup strategy, UI audit & Completed \\ \hline
Weeks 13--14 & Testing and Pre-Handover & Unit/integration/security/load tests, mobile fixes, handover guide, API reference, repository cleanup, CI workflows & Completed \\ \hline
Weeks 15--16 & UAT and Staging Validation & Clean deploy simulation, production verification, UAT persona runs, feedback loop and polish pass & Completed \\ \hline
Weeks 17--18 & Reporting and Submission & Technical report artifacts, screenshots, presentation, final submission hygiene and release tagging & Completed \\ \hline
\bottomrule
\end{longtable}

\subsection{Roles and Responsibilities}

The following table summarizes the major points relevant to this subsection for quick review.


\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{|p{0.26\textwidth}|X|}
\toprule
\textbf{Role Area} & \textbf{Responsibilities} \\ \hline
\midrule
Backend Engineering & API design, schema validation, business logic, role authorization, lifecycle transitions, notification hooks \\ \hline
Frontend Engineering & Route architecture, role dashboards, analyze/result flow, user session handling, responsive behavior \\ \hline
AI and Classification & Vision model integration, rule engine tuning, reasoning fallback, draft generation quality checks \\ \hline
Data and Persistence & MongoDB collection design, indexes, query access paths, reset token TTL and audit metadata \\ \hline
Security and Reliability & Input sanitization, upload hardening, CORS and headers, degrade behavior, readiness/liveness semantics \\ \hline
Testing and QA & Unit and integration tests, resilience and security validation, load profiling, UAT scripts \\ \hline
DevOps and Deployment & Compose consolidation, environment normalization, production verification scripts, backup runbooks \\ \hline
Documentation and Handover & API reference, deployment guides, user manual, architecture and report compilation \\ \hline
\bottomrule
\end{tabularx}
\caption{Role Responsibility Matrix (Project Execution)}
\end{table}

\subsection{Group Dependencies}

The project includes cross-cutting dependencies that were managed explicitly:

\begin{itemize}
\item Backend depends on MongoDB availability and correctly configured environment variables.
\item Analyze flow depends on file storage service and Ollama runtime health.
\item Frontend route behavior depends on JWT session state and role claims.
\item Assignment quality depends on worker profile completeness (department, approval, service area, status).
\item Production readiness depends on compose health checks, proxy alignment, and secure secret configuration.
\end{itemize}

Dependency risks were controlled through health endpoints, fallback behaviors, and scripted verification routines.

\section{Project Scheduling (Gantt/PERT Chart)}

\subsection{High-Level Schedule}

The schedule can be interpreted as a four-phase progression:

\begin{enumerate}
\item Foundation and Core Build (Weeks 1--6)
\item Security and NDMC-oriented Feature Hardening (Weeks 7--14)
\item UAT and Operational Validation (Weeks 15--16)
\item Reporting and Submission Closure (Weeks 17--18)
\end{enumerate}

\subsection{PERT-Like Dependency Flow}

The following table summarizes the major points relevant to this subsection for quick review.


\begin{table}[H]
\centering
\begin{tabular}{|p{0.15\textwidth}|p{0.44\textwidth}|p{0.30\textwidth}|}
\toprule
\textbf{Activity ID} & \textbf{Activity} & \textbf{Depends On} \\ \hline
\midrule
A1 & Core backend and DB setup & None \\ \hline
A2 & Auth and role scaffolding & A1 \\ \hline
A3 & Analyze pipeline integration & A1, A2 \\ \hline
A4 & Frontend role flows & A2, A3 \\ \hline
A5 & Security hardening & A2, A3, A4 \\ \hline
A6 & Assignment, triage, analytics & A3, A4 \\ \hline
A7 & Notification chain and profile features & A4, A6 \\ \hline
A8 & Reliability and load profiling & A5, A6, A7 \\ \hline
A9 & Production deployment profile & A5, A8 \\ \hline
A10 & UAT and acceptance validation & A8, A9 \\ \hline
A11 & Final documentation and submission & A10 \\ \hline
\bottomrule
\end{tabular}
\caption{PERT-Style Activity Dependencies}
\end{table}

\section{Risk Management}

\begin{longtable}{|p{0.24\textwidth}|p{0.15\textwidth}|p{0.15\textwidth}|p{0.38\textwidth}|}
\caption{Risk Register and Mitigation Plan}\\
\toprule
\textbf{Risk} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Mitigation} \\ \hline
\midrule
\endfirsthead
\toprule
\textbf{Risk} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Mitigation} \\ \hline
\midrule
\endhead
Model unavailability or timeout & Medium & High & Multi-tier vision fallback, queue-based generation, retryable 503 responses \\ \hline
Wrong department classification & Medium & High & Deterministic rule engine, triage queue, admin correction workflow \\ \hline
Invalid/malicious uploads & Medium & High & Extension + MIME + magic number + image decode validation, max size enforcement \\ \hline
Unauthorized route access & Low & High & JWT auth, role-gated dependencies, integration tests for permission boundaries \\ \hline
Configuration drift across environments & Medium & Medium & Unified backend environment strategy and scripted deployment checks \\ \hline
Notification inconsistency & Low & Medium & Automated notification-chain tests and mark-all-read verification \\ \hline
Performance degradation under load & Medium & Medium & Profiling, timing logs, load-testing matrix, controlled fallback behavior \\ \hline
Documentation mismatch at handover & Medium & Medium & NDMC deployment guide, API reference sync, verification procedure chapter \\ \hline
\bottomrule
\end{longtable}

\section{Management Outcomes}

The management strategy achieved three notable outcomes:

\begin{itemize}
\item Feature breadth was delivered without compromising security and reliability hardening.
\item Delivery confidence increased through continuous testing and issue-driven iteration.
\item Handover readiness improved through consolidated deployment/configuration artifacts and documentation alignment.
\end{itemize}

\section{Detailed Week-Wise Execution Digest}

\begin{longtable}{|p{0.10\textwidth}|p{0.18\textwidth}|p{0.46\textwidth}|p{0.18\textwidth}|}
\caption{Week-by-Week Execution Summary (W1--W18)}\\
\toprule
\textbf{Week} & \textbf{Date Range} & \textbf{Primary Execution Focus} & \textbf{Management Outcome} \\ \hline
\midrule
\endfirsthead
\toprule
\textbf{Week} & \textbf{Date Range} & \textbf{Primary Execution Focus} & \textbf{Management Outcome} \\ \hline
\midrule
\endhead
W1 & Jan 28--Jan 31 & Repository initialization, environment setup, architecture notes, backend bootstrap & Foundation controls established \\ \hline
W2 & Feb 02--Feb 07 & Mongo connectivity, schema definitions, early CRUD and upload service & Core API baseline available \\ \hline
W3 & Feb 09--Feb 14 & Authentication hardening and classifier exploration & Security and AI direction validated \\ \hline
W4 & Feb 16--Feb 21 & EXIF/location extraction and reverse-geocoding integration & Analyze pipeline readiness achieved \\ \hline
W5 & Feb 23--Feb 28 & Local model runtime setup and draft generation service & AI drafting integrated \\ \hline
W6 & Mar 02--Mar 07 & Frontend routing, upload/result flow, end-to-end integration & Citizen flow stabilized \\ \hline
W7 & Mar 09--Mar 14 & Map refinement, draft persistence, error boundaries, full walkthroughs & Baseline UX hardening complete \\ \hline
W8 & Mar 16--Mar 21 & Security layer (headers, sanitizer, upload validation, limits, indexes) & Production-risk exposure reduced \\ \hline
W9 & Mar 23--Mar 28 & Assignment engine, status timeline, analytics, SLA and bulk controls & Operational depth achieved \\ \hline
W10 & Mar 30--Apr 04 & Notification chain, email stub, password reset, profile update & Communication workflows closed \\ \hline
W11 & Apr 06--Apr 11 & Performance optimization, graceful degrade, profiling, resilience tests & Reliability confidence improved \\ \hline
W12 & Apr 13--Apr 18 & Production container profile and deployment artifact maturation & Deployment readiness strengthened \\ \hline
W13 & Apr 20--Apr 25 & Test expansion (unit/integration/security/load) and mobile fixes & Quality gates expanded \\ \hline
W14 & Apr 27--Apr 30 & API/docs finalization, cleanup, CI/release prep & Pre-handover alignment completed \\ \hline
W15 & May 04--May 09 & Production verification and clean-machine deployment simulation & Reproducibility confirmed \\ \hline
W16 & May 11--May 16 & UAT persona runs, feedback loop, resilience validation & User-level acceptance confidence gained \\ \hline
W17 & May 18--May 23 & Reporting artifacts, screenshots, user manual, presentation draft & Submission assets consolidated \\ \hline
W18 & May 25--May 27 & Final release hygiene and submission closure & Project lifecycle completed \\ \hline
\bottomrule
\end{longtable}

\section{Project Governance Metrics}

\begin{table}[H]
\centering
\begin{tabularx}{\textwidth}{|p{0.34\textwidth}|X|}
\toprule
\textbf{Metric} & \textbf{Interpretation in This Project} \\ \hline
\midrule
Milestone completion consistency & Planned milestone themes were completed in sequence with controlled scope shifts and documented acceleration notes \\ \hline
Defect turnaround behavior & High-severity defects were fixed with targeted regression verification in the same development cycle \\ \hline
Quality gate maturity & Testing moved from smoke-only to layered validation (security, resilience, load, and UAT) \\ \hline
Deployment reproducibility & Scripts and compose profile reduced dependency on manual environment interpretation \\ \hline
Documentation synchronization & Technical docs, runbooks, and report artifacts were updated as implementation evolved \\ \hline
Operational risk closure rate & Key risk classes (auth, upload security, route drift, and notification consistency) were explicitly addressed \\ \hline
\bottomrule
\end{tabularx}
\caption{Governance-Oriented Delivery Metrics}
\end{table}

\section{Management Retrospective}

Three management lessons were central to delivery quality:

\begin{enumerate}
\item \textbf{Early role modeling reduced rework:} Defining citizen/worker/dept-head/admin pathways early prevented significant UI and authorization redesign later.
\item \textbf{Security and reliability should not be deferred:} Integrating these streams before final release avoided unstable late-stage merges.
\item \textbf{Artifact consolidation is a management task:} Unifying environment and compose artifacts reduced support burden and improved handover confidence.
\end{enumerate}

\section{Chapter Summary}

This chapter demonstrated that the project was not only technically implementable but also deliverable within a controlled timeline and governance framework. Feasibility, dependency management, risk planning, and milestone tracking supported predictable progress toward a production-oriented civic platform.
