% !TeX root = ../main.tex
\chapter{System Requirement Study}

This chapter documents requirement analysis for Jan-Sunwai AI, including the baseline state of existing grievance systems, observed limitations, actor expectations, and formalized functional/non-functional requirements. The study was derived from implementation realities, municipal workflow assumptions, and validation through tests and user journeys [1, 2, 11, 12, 13].

\section{Existing System Overview}

Many conventional civic complaint channels depend on one or more of the following interfaces: physical counters, basic web forms, call center logging, or generic ticketing portals. While these methods can register complaints, they often suffer from weak input quality, manual routing dependency, and low traceability.

Key characteristics of existing systems observed in this problem context:

\begin{itemize}
\item Citizens are expected to manually choose departments, often without sufficient guidance.
\item Complaint narratives are text-heavy and unstructured, leading to inconsistent detail quality.
\item Photo evidence may be accepted but is rarely integrated into routing logic.
\item Status updates are frequently sparse or delayed.
\item Workflow metadata (who changed what and why) is often incomplete for audits.
\end{itemize}

\section{Limitations of the Existing System}

Seven recurring limitations were observed in legacy civic grievance systems: manual authority selection by citizens causing frequent misrouting; inconsistent complaint descriptions that require staff to rewrite or clarify cases before action; weak integration of visual evidence leaving photographs unused during triage; limited role-specific workflows that impose the same interface burden on very different personas; minimal lifecycle traceability making accountability reviews difficult; low transparency in queue behavior eroding citizen trust; and sparse administrative analytics that force planning decisions to rely on intuition rather than data [1, 2, 6].

\section{User Characteristics}

The system is designed for four operational personas with distinct interaction patterns and authorization boundaries.

The system serves four distinct operational personas. \textbf{Citizens} report issues and track outcomes: they upload images, review the AI-generated draft, submit their complaint, and monitor notifications and status updates. \textbf{Workers} resolve assigned field tasks by viewing their queue, managing availability, and marking completions with resolution notes. \textbf{Department Heads} manage departmental flow by filtering the department queue, updating complaint status, adding notes, and transferring or escalating cases when required. \textbf{Administrators} exercise end-to-end governance: approving workers, triaging ambiguous AI classifications, performing bulk operations, exporting reports, and monitoring analytics.

\section{Functional Requirements}

\begin{longtable}{|p{0.12\textwidth}|p{0.30\textwidth}|p{0.48\textwidth}|}
\caption{Functional Requirement Specification}\\
\hline
\textbf{FR ID} & \textbf{Requirement Name} & \textbf{Requirement Description} \\ \hline
\hline
\endfirsthead
\hline
\textbf{FR ID} & \textbf{Requirement Name} & \textbf{Requirement Description} \\ \hline
\hline
\endhead
FR-01 & User Registration & System shall allow self-registration for citizen and worker roles with validation and duplicate checks. \\ \hline
FR-02 & User Authentication & System shall establish secure cookie-session authentication on successful login and protect authorized routes, with JWT compatibility for API clients. \\ \hline
FR-03 & Profile Management & Authenticated users shall update selected profile fields such as full name and phone number. \\ \hline
FR-04 & Password Recovery & System shall support forgot-password and reset-password flows with time-bound tokens. \\ \hline
FR-05 & Image Upload Validation & Analyze endpoint shall accept valid JPG/PNG/WEBP images and reject unsupported or malformed payloads. \\ \hline
FR-06 & AI-Assisted Analysis & System shall analyze uploaded image and return category, confidence, location metadata, and draft status. \\ \hline
FR-07 & Draft Generation & System shall produce a complaint draft synchronously or via queued generation, with polling support. \\ \hline
FR-08 & Complaint Creation & System shall create a complaint record from reviewed analysis payload and user-provided description edits. \\ \hline
FR-09 & Complaint Listing & System shall provide role-filtered complaint lists with pagination and optional status filters. \\ \hline
FR-10 & Complaint Detail View & System shall provide complete complaint detail including status history and operational metadata. \\ \hline
FR-11 & Status Update Workflow & Department head/admin shall update complaint status with note and trigger citizen notification. \\ \hline
FR-12 & Department Transfer & Department head/admin shall transfer complaint ownership across departments when misrouted. \\ \hline
FR-13 & Escalation & Department head/admin shall escalate complaints to parent authorities when required. \\ \hline
FR-14 & Worker Assignment & System shall auto-assign eligible workers and support manual override by admin. \\ \hline
FR-15 & Worker Operations & Worker shall manage availability and mark assigned complaints as done. \\ \hline
FR-16 & Notification Center & System shall provide list, unread count, single-read, and mark-all-read operations. \\ \hline
FR-17 & Triage Queue & Admin shall review low-confidence complaints and approve/reject/correct labels. \\ \hline
FR-18 & Analytics & Admin/dept-head shall access overview and geospatial heatmap summaries. \\ \hline
FR-19 & CSV Export & Admin shall export complaint data to CSV for external reporting. \\ \hline
FR-20 & Public Transparency Feed & System shall expose anonymized complaint records without authentication. \\ \hline
FR-21 & Health Monitoring & System shall provide live, ready, model, and GPU health endpoints. \\ \hline
FR-22 & API Version Alias & System shall expose major endpoints under /api/v1 for integration stability. \\ \hline
\hline
\end{longtable}

\section{Non-Functional Requirements}

\begin{longtable}{|p{0.12\textwidth}|p{0.30\textwidth}|p{0.48\textwidth}|}
\caption{Non-Functional Requirement Specification}\\
\hline
\textbf{NFR ID} & \textbf{Quality Attribute} & \textbf{Requirement Description} \\ \hline
\hline
\endfirsthead
\hline
\textbf{NFR ID} & \textbf{Quality Attribute} & \textbf{Requirement Description} \\ \hline
\hline
\endhead
NFR-01 & Security & Cookie-session auth (with JWT compatibility), RBAC, sanitization, secure headers, and account-safe reset flow shall be enforced. \\ \hline
NFR-02 & Reliability & API shall degrade gracefully under model or DB faults with explicit status semantics. \\ \hline
NFR-03 & Performance & Public and health endpoints shall respond consistently under mixed read-heavy load. \\ \hline
NFR-04 & Usability & Role-specific UI shall reduce cognitive load and support responsive layouts for desktop/mobile usage. \\ \hline
NFR-05 & Maintainability & Modular routers/services with documented runbooks shall support iterative enhancement. \\ \hline
NFR-06 & Observability & Process-time headers, rotating logs, and health routes shall support debugging and operations. \\ \hline
NFR-07 & Scalability Readiness & Queue-based generation and container deployment shall support controlled scaling. \\ \hline
NFR-08 & Portability & Project setup shall support Windows and Linux via scripts and Docker-based workflow. \\ \hline
NFR-09 & Compliance Readiness & Local-first inference and sanitized logs shall support stricter governance deployments. \\ \hline
NFR-10 & Data Integrity & Validation schemas and controlled update flows shall minimize malformed or inconsistent records. \\ \hline
\hline
\end{longtable}

\section{Hardware and Software Requirements}

\subsection{Hardware Requirements}

The baseline parameters established below define the minimum viable constraints required to support the proposed features.


\begin{table}[H]
\centering
\caption{Hardware Requirement Baseline}
\begin{tabularx}{\textwidth}{|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|}
\hline
\textbf{Component} & \textbf{Minimum} & \textbf{Recommended} \\ \hline
\hline
CPU & 4-core modern processor & 6 to 8-core processor \\ \hline
System RAM & 12 GB available & 16 GB or higher \\ \hline
GPU VRAM & 4 GB & 6 GB or higher \\ \hline
Storage & 10 GB free & 20 GB free \\ \hline
Network & Stable LAN/internet for maps and package fetch & Managed institutional network \\ \hline
\end{tabularx}
\end{table}

\subsection{Software Requirements}

The baseline parameters established below define the minimum viable constraints required to support the proposed features [16, 11].


\begin{table}[H]
\centering
\caption{Software Environment Requirements}
\begin{tabularx}{\textwidth}{|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|}
\hline
\textbf{Software Component} & \textbf{Baseline} & \textbf{Purpose} \\ \hline
\hline
Operating System & Windows 10+/Linux & Development and deployment host \\ \hline
Python Runtime & Python 3.10+ & Backend and automation scripts \\ \hline
Node.js and npm & Current LTS range & Frontend build and tooling \\ \hline
MongoDB & 7.x class deployment & Document persistence \\ \hline
Ollama Runtime & Local model runtime & Vision/reasoning generation pipeline \\ \hline
Docker Engine/Compose & Docker 24+ recommended & Containerized orchestration \\ \hline
Browser & Chromium/Firefox class & Frontend access and UAT execution \\ \hline
\end{tabularx}
\end{table}

\section{Constraints}

Constraints in this project are not only technical limits but also governance and usability boundaries that directly influence system behavior. The platform must operate under security, role-separation, and infrastructure limitations common in civic environments, while still maintaining user clarity for citizens and workflow efficiency for officials. The following constraint groups define these implementation boundaries and informed several architecture and UI decisions.

\subsection{UI Constraints}

UI constraints define the rendering, state-management, and responsiveness requirements that guard against poor citizen experience in adverse conditions:


\begin{itemize}
\item Role-specific pages must enforce access boundaries and clear redirection behavior.
\item Analyze/result screens must preserve draft continuity during refresh and error conditions.
\item Dashboard cards and maps must remain usable at mobile widths (e.g., 375 px baseline tests).
\end{itemize}

\subsection{Communication Interface}

The platform communicates exclusively over authenticated REST endpoints; the following interface rules govern how clients and services exchange data:


\begin{itemize}
\item REST over HTTP(S) is the canonical communication model.
\item Protected routes require authenticated cookie session or compatibility bearer token.
\item Production frontend proxy must route API traffic to backend service with /api/v1 compatibility.
\end{itemize}

\subsection{Hardware Interface}

Hardware interface constraints bound the AI runtime's operational envelope and determine how the system must gracefully degrade under resource pressure [17]:


\begin{itemize}
\item AI response times depend on model runtime health and available VRAM.
\item GPU contention and model load/unload latency can affect throughput.
\item Large image uploads require strict file-size and decode constraints to avoid resource exhaustion.
\end{itemize}

\subsection{Criticality of Application}

The system is governance-sensitive. Incorrect status transitions, misrouted complaints, or unauthorized updates can directly impact citizen trust and administrative accountability. Hence, status history and role checks are non-negotiable controls [6, 23, 7].

\subsection{Safety and Security Considerations}

Because civic complaints can contain sensitive personal information, the following security constraints are treated as non-negotiable throughout the codebase [7, 8, 9]:


\begin{itemize}
\item Input sanitization is required for free-text fields.
\item File intake must detect extension-header mismatches and corrupted payloads.
\item Token reset flow must avoid account enumeration and raw token persistence.
\item Security headers and CORS policies must remain active on both normal and error responses.
\end{itemize}

\section{Assumptions and Dependencies}

\begin{table}[H]
\centering
\caption{Assumptions and External Dependencies}
\begin{tabularx}{\textwidth}{|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|}
\hline
\textbf{ID} & \textbf{Assumption/Dependency} & \textbf{Implication} \\ \hline
\hline
A1 & MongoDB instance is reachable and healthy & Core read/write workflows remain functional \\ \hline
A2 & Ollama runtime is available with configured models & Analyze and generation pathways remain active \\ \hline
A3 & Environment variables are configured accurately & Auth, DB, CORS, and model endpoints behave as intended \\ \hline
A4 & Frontend and backend URL mapping is consistent & API calls and route protection remain stable \\ \hline
A5 & Worker profiles include valid department and status metadata & Assignment service can select candidates correctly \\ \hline
A6 & Deployment host supports Docker where container mode is used & Production-style stack can be reproduced \\ \hline
\end{tabularx}
\end{table}

\section{Detailed Use-Case Narratives}

\begin{longtable}{|p{0.12\textwidth}|p{0.18\textwidth}|p{0.30\textwidth}|p{0.30\textwidth}|}
\caption{Narrative Use-Case Requirements}\\
\hline
\textbf{UC ID} & \textbf{Actor} & \textbf{Main Flow} & \textbf{Alternate/Exception Flow} \\ \hline
\hline
\endfirsthead
\hline
\textbf{UC ID} & \textbf{Actor} & \textbf{Main Flow} & \textbf{Alternate/Exception Flow} \\ \hline
\hline
\endhead
UC-01 & Citizen & Register account, login, access dashboard & Duplicate username/email is rejected with validation message \\ \hline
UC-02 & Citizen & Upload image and run analyze pipeline & Unsupported/corrupt image rejected with explicit reason \\ \hline
UC-03 & Citizen & Edit draft and submit complaint & If analyze unavailable, user submits manual complaint text after retry guidance \\ \hline
UC-04 & Citizen & Track status and receive notifications & If no notifications available, panel remains stable with unread count zero \\ \hline
UC-05 & Citizen & Submit feedback after complaint resolution & Duplicate feedback attempt blocked with conflict response \\ \hline
UC-06 & Worker & View assigned complaints and update availability & Worker pending approval cannot access worker-only operations \\ \hline
UC-07 & Worker & Mark complaint done & If assignment mismatch occurs, operation is denied by role/ownership checks \\ \hline
UC-08 & Dept Head & Filter department complaints and update status & Dept-head cannot update complaints outside own department \\ \hline
UC-09 & Dept Head & Add notes and transfer/escalate complaint & Invalid complaint ID or missing target yields controlled validation error \\ \hline
UC-10 & Admin & Approve/reject worker registrations & Already-approved/rejected states handled idempotently \\ \hline
UC-11 & Admin & Operate triage queue and analytics & Empty triage queue still returns successful response \\ \hline
UC-12 & Public User & View anonymized complaint board & No authenticated personal fields are exposed in response payload \\ \hline
\end{longtable}

\section{Requirement Priority Classification}

Requirements are classified using the MoSCoW methodology [20] to ensure delivery focus.

\begin{table}[H]
\centering
\caption{MoSCoW-Style Requirement Priority Mapping}
\begin{tabularx}{\textwidth}{|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|}
\hline
\textbf{Priority} & \textbf{Requirement IDs} & \textbf{Interpretation} \\ \hline
\hline
Must Have & FR-01 to FR-12, FR-14, FR-16, FR-21 & Core operational viability: authentication, complaint lifecycle, role control, assignment, notifications, and service health \\ \hline
Should Have & FR-13, FR-15, FR-17, FR-18, FR-22 & Strong governance and maintainability improvements that meaningfully increase institutional readiness \\ \hline
Could Have & FR-19, FR-20 & Non-blocking but high-value transparency and reporting improvements \\ \hline
Non-Functional Critical & NFR-01, NFR-02, NFR-06, NFR-10 & Security, reliability, observability, and data integrity controls required for trusted operation \\ \hline
\end{tabularx}
\end{table}

\section{Chapter Summary}

The requirement study establishes the transition from a manually intensive grievance flow to a structured, secure, and auditable digital workflow. Functional and non-functional requirements were formalized in implementation-aligned terms and mapped against realistic user roles and operational constraints.

