% !TeX root = ../main.tex
\chapter{Limitations and Future Scope}

No production-oriented civic platform is complete without explicit recognition of current boundaries. Jan-Sunwai AI delivers a strong operational baseline, but several limitations remain due to time, infrastructure, and deployment-context constraints [16, 7, 17].

\section{Current Limitations}

Current limitations were documented to provide a realistic boundary for what the present release can guarantee in production-like environments. These limitations do not invalidate the solution; rather, they identify the most important areas where additional engineering and policy integration can strengthen long-term adoption.

\subsection{Model and AI Limitations}

The following AI-specific constraints reflect the boundaries of local model behaviour under diverse civic imagery and describe where the current pipeline still falls short [3, 4, 5]:


\begin{itemize}
\item Classification quality depends on local model behavior and image quality.
\item Ambiguous civic scenes may still require administrative triage despite hybrid logic.
\item Domain adaptation is limited to existing taxonomy and keyword/rule coverage.
\item Draft generation quality can vary by language complexity and scene context.
\end{itemize}

\subsection{Infrastructure Limitations}

Infrastructure ceilings were accepted as deliberate scope boundaries for a project-grade pilot; the specific hardware and queue constraints are [16, 17]:


\begin{itemize}
\item Throughput is constrained by local hardware, especially VRAM and CPU availability.
\item Single-node deployment is suitable for pilot and controlled workloads, not large multi-zone production scale without additional orchestration.
\item Current queue strategy is in-process and may require distributed message infrastructure for higher concurrency targets.
\end{itemize}

\subsection{Workflow and Data Limitations}

Workflow limitations arise from the intersection of civic process complexity and the data quality of administrator-maintained configuration:


\begin{itemize}
\item Worker service-area quality depends on correctness and completeness of admin-maintained profiles.
\item Escalation semantics are configurable but not yet integrated with external authority APIs.
\item Public feed is intentionally anonymized, limiting deeper citizen-facing analytics.
\item Historical trend analytics can be expanded with richer time-series and SLA breach segmentation.
\end{itemize}

\subsection{Security and Governance Limitations}

Security boundaries that fall outside the application layer are inherited from the host deployment environment and must be addressed before large-scale rollout:

\begin{itemize}
\item Security checks are strong at application layer but external perimeter controls (WAF, zero-trust segmentation) are deployment-dependent.
\item Email delivery remains stub-based until institutional SMTP integration is finalized.
\item Continuous dependency vulnerability tracking is not yet fully automated in release workflow.
\end{itemize}

\subsection{Real-World Deployment Challenges}

Each of the following real-world challenges represents an organizational or policy constraint that software engineering alone cannot fully resolve [1, 2, 6, 7]:

\begin{itemize}
\item \textbf{Data Privacy Concerns:} Handling complaints containing sensitive citizen data or private locations requires strict redaction and retention policies.
\item \textbf{Model Bias:} Local models may demonstrate bias if training data lacks representation of diverse civic environments and informal settlements.
\item \textbf{Infrastructure Limitations:} Municipal offices often lack the continuous power supply and network bandwidth required for high-availability AI services.
\item \textbf{Staff Training Requirements:} Shifting from manual data entry to AI-assisted triage demands new digital literacy skills from existing municipal workers.
\end{itemize}

\section{Future Scope}

Future scope is framed as a phased enhancement roadmap that balances impact, effort, and institutional feasibility. The intent is to guide post-submission development toward measurable improvements in intelligence quality, operational scalability, and governance robustness [5, 6, 11, 12].

\subsection{AI and Intelligence Enhancements}

The following enhancements are direct extensions of the existing hybrid classification pipeline and can be built incrementally on top of the current codebase without architectural rewrites:

\begin{enumerate}
\item \textbf{Domain-Adapted Classifier Tuning:} The existing rule engine (already in production) can be extended with a lightweight ONNX classifier trained on municipal complaint image datasets, replacing the current zero-shot vision model for common civic categories.
\item \textbf{Duplicate Complaint Detection:} The complaint MongoDB collection already stores text and category. Adding embedding-based similarity search (via a local embedding model in Ollama) would enable real-time deduplication against the last 30 days of records.
\item \textbf{Multilingual Draft Generation:} The Ollama inference pipeline already supports prompt templating. Switching the draft generation prompt to a multilingual instruction template would unlock Hindi and other regional language output without model changes.
\item \textbf{Confidence-Adaptive Triage Routing:} The current triage queue uses a fixed 0.70 threshold. A simple feedback loop recording admin correction decisions could be used to auto-tune this per-category threshold over time.
\end{enumerate}

\subsection{Operational Workflow Enhancements}

Each workflow enhancement below addresses a specific gap between the current pilot implementation and a full institutional deployment:


\begin{enumerate}
\item Implement SLA policy engine with configurable thresholds per department and issue class.
\item Add escalation automation with conditional approvals and supervisor override logs.
\item Extend assignment with route optimization and field workload balancing by travel cost.
\item Introduce richer citizen feedback loops and resolution quality scoring.
\end{enumerate}

\subsection{Platform and Deployment Enhancements}

Platform and infrastructure improvements address the scale gap between the current Docker Compose pilot and a multi-ward municipal deployment:


\begin{enumerate}
\item Move from single-node compose to orchestrated deployments for higher availability.
\item Add asynchronous message broker for queue resilience and horizontal worker scaling.
\item Integrate centralized observability stack for metrics, traces, and security audit events.
\item Implement backup verification drills and disaster recovery runbooks with recovery point objectives.
\end{enumerate}

\subsection{Security and Compliance Enhancements}

Security and compliance maturity goals represent the delta between the current application-layer hardening and institutional audit readiness [7, 8, 9]:


\begin{enumerate}
\item Introduce periodic penetration testing workflow with report tracking.
\item Add secret rotation automation and hardware-backed key management integration.
\item Expand privacy controls for retention and redaction policies aligned with institutional mandates.
\item Include compliance templates for municipal procurement and audit reporting.
\end{enumerate}

\section{Roadmap Perspective}

\begin{table}[H]
\centering
\caption{Indicative Post-Submission Roadmap}
\begin{tabularx}{\textwidth}{|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|>{\RaggedRight\arraybackslash}X|}
\hline
\textbf{Horizon} & \textbf{Priority Area} & \textbf{Expected Deliverables} \\ \hline
\hline
0--3 months & Stabilization and hardening & SMTP integration, expanded tests, deployment checklist closure \\ \hline
3--6 months & AI and workflow maturity & Better triage intelligence, SLA policy engine, richer analytics \\ \hline
6--12 months & Scale and institutionalization & Distributed queueing, observability stack, multi-zone deployment model \\ \hline
12+ months & Advanced governance intelligence & Predictive demand analytics and department workload forecasting \\ \hline
\end{tabularx}
\end{table}

\section{Risk-Prioritized Improvement Backlog}

\begin{longtable}{|p{0.10\textwidth}|p{0.30\textwidth}|p{0.13\textwidth}|p{0.13\textwidth}|p{0.24\textwidth}|}
\caption{Future Backlog Prioritization Matrix}\\
\hline
\textbf{ID} & \textbf{Improvement Item} & \textbf{Impact} & \textbf{Effort} & \textbf{Suggested Horizon} \\ \hline
\hline
\endfirsthead
\hline
\textbf{ID} & \textbf{Improvement Item} & \textbf{Impact} & \textbf{Effort} & \textbf{Suggested Horizon} \\ \hline
\hline
\endhead
FB-01 & SMTP relay production integration with delivery audit logs & High & Medium & 0--3 months \\ \hline
FB-02 & Centralized secrets management and rotation policy & High & Medium & 0--6 months \\ \hline
FB-03 & Distributed queue broker for LLM jobs & High & High & 3--9 months \\ \hline
FB-04 & Expanded language support in generated drafts & Medium & Medium & 3--6 months \\ \hline
FB-05 & Domain-adapted classifier tuning dataset & High & High & 6--12 months \\ \hline
FB-06 & Duplicate complaint detection using embeddings & Medium & High & 6--12 months \\ \hline
FB-07 & SLA rules engine with department-specific thresholds & High & Medium & 3--9 months \\ \hline
FB-08 & Escalation policy simulator for governance review & Medium & Medium & 6--12 months \\ \hline
FB-09 & Citizen voice intake and transcription & Medium & High & 9--15 months \\ \hline
FB-10 & Geo-aware workload balancing with travel-time estimation & Medium & High & 6--12 months \\ \hline
FB-11 & Real-time websocket notification stream & Medium & Medium & 3--6 months \\ \hline
FB-12 & Expanded public analytics visualizations & Medium & Medium & 3--9 months \\ \hline
FB-13 & Multi-tenant authority hierarchy support & High & High & 9--18 months \\ \hline
FB-14 & Formal accessibility conformance audit (WCAG) & High & Medium & 0--6 months \\ \hline
FB-15 & Integrated vulnerability scanning in CI pipeline & High & Low & 0--3 months \\ \hline
FB-16 & Observability stack with trace correlation & High & Medium & 3--9 months \\ \hline
FB-17 & Offline-first synchronization for field staff app mode & Medium & High & 9--18 months \\ \hline
FB-18 & Complaint closure quality scoring and follow-up loop & Medium & Medium & 3--9 months \\ \hline
FB-19 & Auto-generated administrative weekly digest reports & Low & Low & 0--3 months \\ \hline
FB-20 & Predictive hotspot forecasting for civic incidents & Medium & High & 12+ months \\ \hline
\hline
\end{longtable}

\section{Adoption Considerations for Municipal Deployment}

Municipal adoption requires alignment between software capability, administrative processes, and organizational readiness. This section therefore highlights the non-technical conditions that significantly influence rollout success even when the technical platform is stable.

\subsection{Institutional Readiness Factors}

Municipal deployment success is influenced by organizational readiness as much as software maturity. Key factors include role training quality, escalation policy clarity, and data-governance ownership. Even technically complete systems can underperform if operational owners do not have clear process responsibility.

\subsection{Change Management Considerations}

Key considerations include:


\begin{itemize}
\item Staff onboarding must include role-specific workflow rehearsals, not only UI tours.
\item Escalation policy language should be standardized before production launch.
\item Existing manual grievance channels should run in parallel during pilot transition.
\item Citizen communication templates should be localized and reviewed for clarity.
\end{itemize}

\subsection{Data Governance Considerations}

Future rollout should include an explicit policy for retention, redaction, access audits, and report publication limits. This is particularly important where complaint evidence may include personal or location-sensitive context.

\section{Research Continuation Opportunities}

The platform also creates opportunities for academic extension:

\begin{enumerate}
\item Comparative analysis of hybrid rule+LLM pipelines versus end-to-end multimodal classifiers in civic datasets.
\item Measurement of complaint resolution outcomes before and after AI-assisted intake.
\item Bias and fairness analysis across department categories and geographies.
\item Human-in-the-loop triage ergonomics and decision consistency studies.
\end{enumerate}

\section{Chapter Summary}

The current system is robust for pilot and structured deployment but remains intentionally conservative in distributed scaling and advanced automation. The future scope presented here provides a realistic pathway from project-grade implementation to institutional-grade civic infrastructure.
