% !TeX root = ../main.tex
\chapter{Appendices}

\section{Appendix A: API Endpoint Snapshot}

This appendix summarizes the primary endpoint inventory implemented in the current release. All major endpoint groups are additionally exposed under the \texttt{/api/v1} alias path.

\small
\begin{itemize}[leftmargin=1.5em, noitemsep]
    \item \textbf{POST} \path{/api/v1/users/register}: Citizen/worker self-registration
    \item \textbf{POST} \path{/api/v1/users/login}: Cookie-session login
    \item \textbf{GET} \path{/api/v1/users/me}: Current user profile
    \item \textbf{POST} \path{/api/v1/analyze}: Image analysis and draft initiation
    \item \textbf{POST} \path{/api/v1/complaints}: Create complaint
    \item \textbf{GET} \path{/api/v1/complaints}: Role-filtered complaint list
    \item \textbf{PATCH} \path{/api/v1/complaints/{id}/status}: Status transition (Admin/Dept Head)
    \item \textbf{PATCH} \path{/api/v1/complaints/{id}/transfer}: Department transfer
    \item \textbf{PATCH} \path{/workers/me/status}: Availability update (Worker)
    \item \textbf{PATCH} \path{/workers/{id}/approve}: Approve worker (Admin)
    \item \textbf{GET} \path{/api/v1/notifications}: User notification list
    \item \textbf{GET} \path{/triage/review-queue}: Low-confidence triage list (Admin)
    \item \textbf{GET} \path{/api/v1/analytics/overview}: High-level operational metrics
    \item \textbf{GET} \path{/api/v1/health/live|ready}: Service health probes
\end{itemize}
\normalsize

\section{Appendix B: Environment Variable Catalog}

\begin{itemize}[leftmargin=1.5em, noitemsep]
    \item \path{MONGODB_URL}: Primary Mongo connection URI.
    \item \path{DB_NAME}: Target database name (e.g., \path{jan_sunwai_db}).
    \item \path{JWT_SECRET_KEY}: Key used for session token signing.
    \item \path{OLLAMA_BASE_URL}: Endpoint for the local model runtime.
    \item \path{VISION_MODEL}: Primary model for image extraction (e.g., \path{qwen2.5vl}).
    \item \path{REASONING_MODEL}: Model used for drafting (e.g., \path{llama3.2}).
    \item \path{AMBIGUITY_THRESHOLD}: Logic boundary for triage routing.
    \item \path{ALLOWED_ORIGINS}: CORS policy for frontend dashboard access.
\end{itemize}
\normalsize

\section{Appendix C: Department Taxonomy and Authority Routing}

\begin{description}
\item[Health Department:] MUNI\_SANITATION $\rightarrow$ Escalates to STATE\_CMO
\item[Civil Department:] MUNI\_PWD $\rightarrow$ Escalates to STATE\_CMO
\item[Horticulture:] MUNI\_HORTICULTURE $\rightarrow$ Escalates to STATE\_CMO
\item[Electrical Department:] UTIL\_DISCOM $\rightarrow$ Escalates to STATE\_CMO
\item[IT Department:] STATE\_CMO $\rightarrow$ Escalates to CENTRAL\_DARPG
\item[Commercial:] STATE\_CMO $\rightarrow$ Escalates to CENTRAL\_DARPG
\item[Enforcement:] POLICE\_TRAFFIC $\rightarrow$ Escalates to STATE\_CMO
\item[Fire Department:] POLICE\_LOCAL $\rightarrow$ Escalates to STATE\_CMO
\item[Uncategorized:] Rule/keyword fallback $\rightarrow$ Escalates to Nullable
\end{description}

\section{Appendix D: Requirement-to-Test Traceability Matrix}

\small
\begin{description}
\item[FR-01 (User registration flow):] Register route tests and manual registration UI checks (Verified).
\item[FR-02 (Login and cookie-session issuance):] Auth permission tests and integration smoke login path (Verified).
\item[FR-03 (Profile editing):] API endpoint behavior and profile page walkthrough (Verified).
\item[FR-04 (Password recovery/reset):] Forgot/reset endpoint implementation and UI path checks (Verified).
\item[FR-05 (Upload validation):] Resilience/security tests for header mismatch and size limits (Verified).
\item[FR-06 (AI-assisted analysis):] Analyze endpoint response contract checks and manual scenario runs (Verified).
\item[FR-07 (Draft regeneration queue):] Generation polling route and regeneration request flow (Verified).
\item[FR-08 (Complaint creation):] End-to-end citizen flow and complaint list visibility checks (Verified).
\item[FR-09 (Role-filtered listing):] Role-based list query behavior and dashboard checks (Verified).
\item[FR-10 (Complaint detail):] Complaint detail retrieval and timeline rendering checks (Verified).
\item[FR-11 (Status update workflow):] Notification chain tests and status history assertions (Verified).
\item[FR-12 (Department transfer):] Admin/dept-head transfer operation checks (Verified).
\item[FR-13 (Escalation support):] Escalation endpoint and timeline update behavior checks (Verified).
\item[FR-14 (Worker assignment):] Assignment service behavior and admin assignment controls (Verified).
\item[FR-15 (Worker task completion):] Worker done workflow and slot-free behavior checks (Verified).
\item[FR-16 (Notification center):] Unread-count and mark-read chain tests (Verified).
\item[FR-17 (Triage queue governance):] Triage review UI and decision endpoint flow checks (Verified).
\item[FR-18 (Analytics and heatmap):] Admin analytics and map rendering walkthrough (Verified).
\item[FR-19 (CSV export):] Export endpoint execution and output schema checks (Verified).
\item[FR-20 (Public board):] Public feed route and UI accessibility checks (Verified).
\item[FR-21 (Health probes):] Live, ready, model, and GPU endpoint smoke checks (Verified).
\item[FR-22 (API version alias):] /api/v1 integration checks in matrix tests (Verified).
\end{description}
\normalsize
