\chapter{Report Verification Procedure}

This chapter defines a structured verification process to ensure that the final report and associated project artifacts are consistent, reproducible, and suitable for academic and technical evaluation.

\section{Verification Objectives}

\begin{itemize}
\item Verify that the report accurately reflects the implemented system.
\item Validate consistency between codebase features, documentation, and chapter claims.
\item Confirm reproducibility of setup, execution, and testing steps.
\item Ensure chapter formatting and artifact completeness for final submission.
\end{itemize}

\section{Document Verification Checklist}

\begin{longtable}{|p{0.08\textwidth}|p{0.58\textwidth}|p{0.24\textwidth}|}
\caption{Document-Level Verification Checklist}
\toprule
\textbf{ID} & \textbf{Verification Item} & \textbf{Status (Y/N/NA)} \\
\midrule
\endfirsthead
\toprule
\textbf{ID} & \textbf{Verification Item} & \textbf{Status (Y/N/NA)} \\
\midrule
\endhead
DV-01 & Title page contains project identity, session, and submission context \\
DV-02 & Table of contents matches chapter order and section numbering \\
DV-03 & Lists of figures/tables/abbreviations/definitions are present \\
DV-04 & All eleven required chapters are present in sequence \\
DV-05 & Figures and tables include captions and are referenced in context \\
DV-06 & Screenshot list is aligned with required checklist items \\
DV-07 & Literature and references are cited in proper academic style \\
DV-08 & No contradictory statements across chapters and appendices \\
DV-09 & Grammar and formatting consistency (font, spacing, headings) is maintained \\
DV-10 & Final PDF metadata and hyperlinks are functioning \\
\bottomrule
\end{longtable}

\section{Technical Verification Checklist}

\begin{longtable}{|p{0.08\textwidth}|p{0.58\textwidth}|p{0.24\textwidth}|}
\caption{Technical Consistency Verification Checklist}
\toprule
\textbf{ID} & \textbf{Verification Item} & \textbf{Status (Y/N/NA)} \\
\midrule
\endfirsthead
\toprule
\textbf{ID} & \textbf{Verification Item} & \textbf{Status (Y/N/NA)} \\
\midrule
\endhead
TV-01 & API endpoint inventory in report matches implemented routers \\
TV-02 & Role definitions in report match auth and dashboard behavior \\
TV-03 & Database collections and indexes described match code-level setup \\
TV-04 & Security controls listed are present in middleware/services/tests \\
TV-05 & Deployment architecture section matches compose and Docker artifacts \\
TV-06 & Testing chapter reflects actual test suites and scripts \\
TV-07 & UAT narrative aligns with documented UAT plan and acceptance checks \\
TV-08 & Known limitations section reflects current implementation boundaries \\
TV-09 & Future scope remains realistic relative to current architecture \\
TV-10 & Appendices contain accurate and non-contradictory operational details \\
\bottomrule
\end{longtable}

\section{Reproducibility Procedure}

\subsection{Build and Run Verification}

\begin{enumerate}
\item Verify backend environment file is configured.
\item Start MongoDB dependency using Docker Compose.
\item Start backend service and confirm live and ready probes.
\item Start frontend service and confirm route accessibility.
\item Execute a sample citizen complaint flow from upload to submission.
\item Validate status update and notification chain behavior.
\end{enumerate}

\subsection{Test Verification}

\begin{enumerate}
\item Execute backend automated tests through provided script wrapper.
\item Execute security and resilience test scripts.
\item Run notification chain test script.
\item Execute load test in baseline profile and collect key metrics.
\item Confirm no critical failures in logs and test outputs.
\end{enumerate}

\subsection{Deployment Verification}

\begin{enumerate}
\item Bring up compose stack with production profile.
\item Check health status for backend and database services.
\item Validate frontend proxy access to API versioned paths.
\item Confirm upload persistence path correctness in container volume mapping.
\item Validate index creation and backup script execution readiness.
\end{enumerate}

\section{Artifact Completeness Checklist}

\begin{table}[H]
\centering
\caption{Submission Artifact Completeness}
\begin{tabularx}{\textwidth}{|p{0.34\textwidth}|p{0.18\textwidth}|X|}
\toprule
\textbf{Artifact} & \textbf{Available} & \textbf{Notes} \\
\midrule
Source code repository & Yes & Backend, frontend, scripts, docs \\
Final report LaTeX source & Yes & Chapter-wise modular structure \\
Compiled final report PDF & To be generated & Build from main.tex \\
API reference document & Yes & Project documentation included \\
Deployment guide & Yes & NDMC deployment and production verification guides \\
Security/load/UAT runbooks & Yes & Dedicated documentation and scripts \\
Screenshot pack & Partial/To update & Placeholders prepared; final captures can be inserted \\
Presentation artifact & External & To be attached as submission companion \\
\bottomrule
\end{tabularx}
\end{table}

\section{Final Sign-Off Template}

\noindent
\textbf{Verifier Name:} \rule{0.55\textwidth}{0.4pt}\\[0.6em]
\textbf{Role:} \rule{0.69\textwidth}{0.4pt}\\[0.6em]
\textbf{Verification Date:} \rule{0.53\textwidth}{0.4pt}\\[0.6em]
\textbf{Overall Status:} \rule{0.56\textwidth}{0.4pt}\\[0.6em]
\textbf{Remarks:}\\[0.4em]
\rule{0.95\textwidth}{0.4pt}\\[1.2em]
\rule{0.95\textwidth}{0.4pt}\\[1.2em]
\rule{0.95\textwidth}{0.4pt}\\[1.6em]
\textbf{Signature:} \rule{0.60\textwidth}{0.4pt}

\section{Evidence Mapping Matrix}

\begin{longtable}{|p{0.16\textwidth}|p{0.30\textwidth}|p{0.18\textwidth}|p{0.28\textwidth}|}
\caption{Report Claim to Artifact Evidence Mapping}
\toprule
\textbf{Claim Area} & \textbf{Evidence Artifact} & \textbf{Evidence Type} & \textbf{Verification Note} \\
\midrule
\endfirsthead
\toprule
\textbf{Claim Area} & \textbf{Evidence Artifact} & \textbf{Evidence Type} & \textbf{Verification Note} \\
\midrule
\endhead
Architecture design & System architecture documentation & Technical doc & Component boundaries and flow definitions match chapter design claims \\
Department taxonomy & Department hierarchy document & Technical doc & Canonical categories align with classifier/routing behavior \\
Schema and indexing & Schema design document and index scripts & Technical doc + code artifact & Collections and indexes align with database design sections \\
API surface & API reference documentation & Technical doc & Endpoint inventory aligns with appendix snapshots \\
Auth and role controls & Auth test suite and role dependency code & Test + code artifact & Role restrictions and claims behavior validated \\
Notification behavior & Notification chain tests and UI screenshots & Test + visual evidence & Badge update and mark-read behavior verified \\
Security hardening & Security testing guide and resilience tests & Test documentation + code tests & Upload validation, sanitization, and headers confirmed \\
Performance and load profile & Load-testing guide and locust scripts & Test runbook + script & Profile definitions and acceptance semantics documented \\
UAT execution & UAT plan and persona scripts & Validation plan & Citizen/admin workflows and acceptance criteria documented \\
Deployment readiness & Deployment and production verification docs & Ops runbook & Compose profile and environment steps reproducible \\
Frontend recovery flow & Login/forgot/reset page screenshots & Visual evidence & Password recovery UX claims traceable \\
Triage governance & Triage queue screenshot and endpoint behavior & Visual + API evidence & Low-confidence review path validated \\
Analytics visibility & Analytics and heatmap screenshots & Visual evidence & Dashboard claims supported by rendered views \\
Public transparency & Public board screenshot & Visual evidence & Anonymous status listing demonstrated \\
Citizen timeline and escalation & Dashboard timeline screenshots & Visual evidence & Status history and escalation note rendering confirmed \\
\bottomrule
\end{longtable}

\section{Submission-Day Verification Script}

\subsection{Pre-Compilation Steps}

\begin{enumerate}
\item Confirm all chapter files are present and referenced in \texttt{main.tex}.
\item Place available screenshots in \texttt{docs/images/screenshots/} using expected names.
\item Update title page institute and contributor details if required by the evaluator.
\item Re-check chapter numbering and heading consistency.
\end{enumerate}

\subsection{Compilation Steps}

\begin{enumerate}
\item Run \texttt{pdflatex -interaction=nonstopmode main.tex}.
\item Run the same command a second time to stabilize TOC, lists, and references.
\item Verify generated PDF for figure/table index completeness.
\item Confirm page range target (70--100) is met and note final page count in submission log.
\end{enumerate}

\subsection{Final QA Steps}

\begin{enumerate}
\item Check screenshot pages for missing placeholders and replace pending files as needed.
\item Validate references, captions, and table readability in print layout.
\item Archive LaTeX source, PDF output, and supporting evidence pack together.
\item Record final verification sign-off with date and reviewer name.
\end{enumerate}

\section{Chapter Summary}

The verification procedure ensures that report claims remain aligned with implementation reality and that submission quality is traceable. By combining document checks, technical consistency checks, reproducibility checks, and sign-off structure, this chapter closes the report with a practical audit trail for evaluation.
