\chapter*{List of Definitions}
\addcontentsline{toc}{chapter}{List of Definitions}
\label{chap:list-definitions}

\begin{longtable}{p{0.30\textwidth}p{0.64\textwidth}}
\midrule
\textbf{Term} & \textbf{Definition} \\ \midrule
\endfirsthead
\midrule
\textbf{Term} & \textbf{Definition} \\ \midrule
\endhead
Civic Complaint & A citizen-reported issue related to municipal infrastructure, utilities, sanitation, public safety, or governance services. \\ \midrule
Complaint Lifecycle & Ordered state progression of a complaint such as Open, In Progress, Resolved, and Rejected. \\ \midrule
Status History & Auditable log of state changes with timestamp, actor identity, and note. \\ \midrule
Authority Routing & Mapping from classified category to responsible authority identifier and escalation parent. \\ \midrule
Triage Queue & Review queue for low-confidence AI outputs requiring administrator decision. \\ \midrule
Service Area & Worker-specific geospatial zone defined by latitude, longitude, and radius in kilometers. \\ \midrule
Auto-Assignment & Rule-based process selecting an eligible worker based on department, availability, and workload. \\ \midrule
Escalation Loop & Background process that pushes unresolved complaints to higher authorities when SLA conditions fail. \\ \midrule
Rule Engine & Deterministic keyword and scoring logic used to improve category stability before optional LLM reasoning. \\ \midrule
Hybrid Pipeline & Multi-stage architecture combining vision extraction, deterministic rules, and selective reasoning model invocation. \\ \midrule
Draft Generation & AI-assisted creation of formal grievance text from image evidence and extracted metadata. \\ \midrule
Graceful Degradation & Controlled failure behavior where the API returns meaningful retryable errors without service crash. \\ \midrule
Magic Number Validation & File integrity check that verifies image binary headers against extension and MIME type. \\ \midrule
Public Transparency Feed & Anonymized read-only endpoint exposing selected complaint records for public visibility. \\ \midrule
Readiness Probe & Endpoint validating dependency availability, especially MongoDB and model runtime health. \\ \midrule
Rate Limiting & Request throttling control to mitigate abuse and resource exhaustion. \\ \midrule
Token Hashing & Storage of reset-token digest instead of raw token to reduce credential leakage risk. \\ \midrule
Canonical Taxonomy & Standardized department labels used across classifier output, dashboards, and routing. \\ \midrule
Operational Governance & Administrative controls for approvals, transfers, notes, audit review, and reporting. \\ \midrule
\end{longtable}
